---
title: "Prediction Assignment"
author: "Sanjay Kumar"
date: "28-Dec-2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.


## Loading Data and Cleaning

You can also embed plots, for example:

```{r, echo=TRUE,cache=TRUE}
#download files
 if (!file.exists("./pml-training.csv")){
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile="./pml-training.csv")
 }
if (!file.exists("./pml-testing.csv")){
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile="./pml-testing.csv")
}
#Load data
training<-read.csv("pml-training.csv")
testing<-read.csv("pml-testing.csv")
dim(training)
dim(testing)
#Remove near zero variance column
library(caret)
training1<-training[, -nearZeroVar(training)]

#mostly null

training<-training1[, -which(colMeans(is.na(training1)) > 0.5)]

#remmove first 7 columns which are username or timestamp
 training<-training[,-c(1:7)]


#testing<-testing1[complete.cases(testing1), ]
dim(training)
#remove null row
training1<-training[complete.cases(training), ]
dim(training1)

```
## Data partition 


```{r, echo=TRUE}
#create data partition for training and test
library(caret)
dpart<-createDataPartition(training1$classe,p=0.7,list=FALSE)
trainSet<-training[dpart,]
testSet<-training[-dpart,]
dim(trainSet)
dim(testSet)

```
## Prediction Model
The model run time exceeds 30 min. Please be patient
```{r, echo=TRUE}
set.seed(1234)
#use 10 fold cross validation
fitControl <- trainControl(method = "cv",number = 10)
#use random forest method
rf<-train(classe~.,method="rf",trControl = fitControl,data=trainSet,verbose = FALSE)
#gradient boost
gbm<-train(classe~.,method="gbm",trControl = fitControl,data=trainSet,verbose = FALSE)
#linear discriminator
lda<-train(classe~.,method="lda",trControl = fitControl,data=trainSet,verbose = FALSE)

predrf<-predict(rf,testSet)
predgbm<-predict(gbm,testSet)
predlda<-predict(lda,testSet)

confusionMatrix(predrf,testSet$classe)
confusionMatrix(predgbm,testSet$classe)
confusionMatrix(predlda,testSet$classe)

#pred<-predict(rf,testSet$classe)

```
## Using model to predict class

```{r, echo=TRUE}
# remove unnecessary columns from testing
#keep<-names(trainSet)
#testing1<-testing[,(names(testing) %in% keep)]
#dim(testing1)
set.seed(1234)
values<-predict(rf,testing)

```
## Conclusion
The final answer is given below which has 99.56% accuracy using random forest method
```{r, echo=FALSE}
values
```